Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-14 csf-2017/paper.tex
update to math
added a couple of comments

% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
\dominique{it is not explained what cofes are?}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  {\small
    \begin{align*}              
      \xi : \Wor & \cong \blater (\nats \finparfun \Regions)\\
      \Regions & = \{\revoked\} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\temp\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
                   \end{multlined} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\perma\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
                   \end{multlined}
    \end{align*}
  } and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
the worlds:
\begin{align*}
  \Worlds & = \RegionNames \finparfun \Regions\\
 \RegionNames & = \nats
\end{align*}

@@

The \emph{public future} region relation satisfies \dominique{is defined by?} the following properties:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
    (s,s') \in \phi_\pub }
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
In the public future regions we do not permit revocation of temporary regions.
Instead, temporary and permanent regions are only allowed to take a transition
in the public part of the transition system. Revoked regions can either remain
revoked or be masked away by a temporary region. This means that the public
future world relations allows us to reuse a region that has been revoked
earlier.

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
  \begin{align*}
    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
    &\qquad\begin{multlined}
      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow 
      \dom(\ms) = \dom(\ms') \text{\footnotesize{ and }} \\
      \forall \addr \in
      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
    \end{multlined}
  \end{align*}

>>>>

% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
 \dominique{it is not explained what cofes are?}
 \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  {\small
    \begin{align*}              
      \xi : \Wor & \cong \blater (\nats \finparfun \Regions)\\
      \Regions & = \{\revoked\} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\temp\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))
                   \end{multlined} \uplus \\
                 &\;\quad \begin{multlined}
                     \{\perma\} \times \States \times \Rels \times \\
                     (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
                   \end{multlined}
    \end{align*}
  } and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
the worlds:
\begin{align*}
  \Worlds & = \RegionNames \finparfun \Regions\\
 \RegionNames & = \nats
\end{align*}

@@

The \emph{public future} region relation satisfies
 \dominique{is defined by?} 
 \lau{As I have understood it, the relation from the theorem defines the future world relations and indirectly the future region relations. It can be shown that these will have the properties mentioned below}
the following properties:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
    (s,s') \in \phi_\pub }
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
In the public future regions we do not permit revocation of temporary regions.
Instead, temporary and permanent regions are only allowed to take a transition
in the public part of the transition system. Revoked regions can either remain
revoked or be masked away by a temporary region. This means that the public
future world relations allows us to reuse a region that has been revoked
earlier.

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{
    \npair{(\start,\addrend)}
    \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; W(r) \text{\footnotesize{ is address-stratified and}} \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r)\nsupsim[n-1] \iota_{\start',\addrend'}
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{
    \begin{multlined}
      \npairP{(\var{perms}, \\
        \start,\addrend)}
    \end{multlined}
     \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W, a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
  \end{align*}
  \begin{align*}
    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
    &\qquad
      \begin{aligned}
        &\forall s', \hat{W}, n, \ms, \ms' \ldotp \\
        &\;\;\npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow\\
        &\;\;\;\;\dom(\ms) = \dom(\ms') \text{\footnotesize{ and }} \\
        &\;\;\;\;\forall \addr \in \dom(\ms)\ldotp\\
        &\;\;\;\;\;\;\npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
      \end{aligned}
  \end{align*}
